---
title: "Premier League 2021/2022 Player Stat Report"
output: 
  html_document:
    theme: united
    highlight: tango
    includes:
      after_body: footer.html
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Load modules required for analysis
library(tidyverse)
library(rvest)
library(ggplot2)
library(ggiraph)
library(ggrepel)
library(scales)
library(pracma)
library(RColorBrewer)
library(fmsb)

# Team names - these are different for different data sources (i.e. ESPN and Understat)
results_team_names = c("Arsenal", "Aston Villa", "Brentford", "Brighton & Hove Albion", "Burnley",
                       "Chelsea", "Crystal Palace", "Everton", "Leeds United", "Leicester City",
                       "Liverpool", "Manchester City", "Manchester United", "Newcastle United",
                       "Norwich City", "Southampton", "Tottenham Hotspur",
                       "Watford", "West Ham United", "Wolverhampton Wanderers")

xG_team_names = c("Arsenal", "Aston Villa", "Brentford", "Brighton", "Burnley", "Chelsea",
                  "Crystal Palace", "Everton", "Leeds", "Leicester",
                  "Liverpool", "Manchester City", "Manchester United", "Newcastle United",
                  "Norwich", "Southampton", "Tottenham", "Watford",
                  "West Ham", "Wolverhampton Wanderers")

# 1. Results analysis

# Scrape results for every Premier league team
source("scrape_results.R")
full_results = scrape_results()

# Calculate home/away win/draw/loss totals and percentages for a PARTICULAR team
source("process_data.R")
source("home_and_away.R")

# Calculate win/draw/loss totals and percentages for home and away games for EVERY team
source("summary_all.R")
processed_results = summary_all(full_results, team_names)

# Calculate home and away rankings
source("home_and_away_rankings.R")
home_rankings <- home_rankings(processed_results, team_names)
away_rankings <- away_rankings(processed_results, team_names)

# Make team names consistent across all dataframes
home_rankings <- replace(home_rankings, home_rankings == "Brighton & Hove Albion", "Brighton")
home_rankings <- replace(home_rankings, home_rankings == "Leeds United", "Leeds")
home_rankings <- replace(home_rankings, home_rankings == "Leicester City", "Leicester")
home_rankings <- replace(home_rankings, home_rankings == "Norwich City", "Norwich")
home_rankings <- replace(home_rankings, home_rankings == "Tottenham Hotspur", "Tottenham")
home_rankings <- replace(home_rankings, home_rankings == "West Ham United", "West Ham")

away_rankings <- replace(away_rankings, away_rankings == "Brighton & Hove Albion", "Brighton")
away_rankings <- replace(away_rankings, away_rankings == "Leeds United", "Leeds")
away_rankings <- replace(away_rankings, away_rankings == "Leicester City", "Leicester")
away_rankings <- replace(away_rankings, away_rankings == "Norwich City", "Norwich")
away_rankings <- replace(away_rankings, away_rankings == "Tottenham Hotspur", "Tottenham")
away_rankings <- replace(away_rankings, away_rankings == "West Ham United", "West Ham")

# 2. Team xG

# Scrape team xG data
source("teams_data.R")
team_xG <- get_team_data()

# Calculate differences between goals/points and expected goals/points
team_xG <- team_xG %>% 
  add_column(Goals_minus_xG = team_xG$Goals - team_xG$xG,
             GA_minus_xGA = team_xG$GA - team_xG$xGA,
             Points_minus_xPTS = team_xG$Points - team_xG$xPTS)

# Create a league ordered by xPTS (expected points)
league_by_xpts <- team_xG %>%
  select(Team, Matches, Wins, Draws, Losses, Goals, GA, Points, xG, xGA, xPTS) %>%
  arrange(-xPTS)


# 3. Player xG

# Scrape player xG data
source("players_data.R")
player_xG <- get_player_data()

# Apostrophes have been encoded strangely - this is corrected
player_xG$player_name <- gsub("&#039;", "'", player_xG$player_name)

# There are two players called Emerson - for uniqueness, we amend these entries
player_xG$player_name[player_xG$player_name == "Emerson" & player_xG$team_title == "Chelsea"] = "Emerson Palmieri"
player_xG$player_name[player_xG$player_name == "Emerson" & player_xG$team_title == "Tottenham"] = "Emerson Royal"
```

***

Data from ESPN.co.uk Premier League results: espn.co.uk/football/ and understat.com Expected goals (xG) and shot data: understat.com/league/EPL accessed on `r format(Sys.Date(), "%d-%m-%Y")`.

```{r, echo=FALSE}
# Select a player
selectInput("player", "Select a player: ",
            choices = player_xG$player_name,
            selected = "Harry Kane",
            multiple = FALSE)

# Determine team of player chosen
team <- reactive({
  player_xG$team_title[player_xG$player_name == input$player]
})
```

`r renderUI({HTML(paste("<b>Player:</b>", input$player, 
  sep = " ", collapse = ", "))})`

`r renderUI({HTML(paste("<b>Team:</b>", team(), 
  sep = " ", collapse = ", "))})`
  
```{r, echo=FALSE}
# Calculate number of wins, draws and losses for the player's team
wins <- reactive({
  team_xG$Wins[team_xG$Team == team()]
}) 

draws <- reactive({
  team_xG$Draws[team_xG$Team == team()]
})

losses <- reactive({
  team_xG$Losses[team_xG$Team == team()]
})
```

```{r, echo=FALSE}
# Calculate home/away win percentage and rankings
home_win <- reactive({
  home_rankings$PERCENTAGE[home_rankings$TEAM == team() & home_rankings$OUTCOME == "WIN"]
})

home_win_ranking <- reactive({
  home_rankings$RANK[home_rankings$TEAM == team() & home_rankings$OUTCOME == "WIN"]
})

away_win <- reactive({
  away_rankings$PERCENTAGE[away_rankings$TEAM == team() & away_rankings$OUTCOME == "WIN"]
})

away_win_ranking <- reactive({
  away_rankings$RANK[away_rankings$TEAM == team() & away_rankings$OUTCOME == "WIN"]
})

```

```{r, echo=FALSE}
# Calculate number of goals, xG and xG delta
goals <- reactive({
  player_xG$goals[player_xG$player_name == input$player]
})

xG <- reactive({
  player_xG$xG[player_xG$player_name == input$player]
})

xG_delta <- reactive({
  goals() - xG()
})

# Add underperforming/overperforming information
player_xG <- player_xG %>%
  mutate(Performance = case_when(
    goals - xG > 0 ~ "overperforming",
    goals - xG < 0 ~ "underperforming",
    goals - xG == 0 ~ "matching"
  ))

player_xG <- player_xG %>%
  mutate(Pos_Neg = case_when(
    goals - xG > 0 ~ "positive",
    goals - xG < 0 ~ "negative",
    goals - xG == 0 ~ "neutral"
  ))

performance <- reactive({
  player_xG$Performance[player_xG$player_name == input$player]
})

pos_neg <- reactive({
  player_xG$Pos_Neg[player_xG$player_name == input$player]
})
```

```{r, echo=FALSE}
# Calculate number of assists, xA and xA delta
assists <- reactive({
  player_xG$assists[player_xG$player_name == input$player]
})

xA <- reactive({
  player_xG$xA[player_xG$player_name == input$player]
})

xA_delta <- reactive({
  assists() - xA()
})

# Add underperforming/overperforming information
player_xG <- player_xG %>%
  mutate(Performance_A = case_when(
    assists - xA > 0 ~ "overperforming",
    assists - xA < 0 ~ "underperforming",
    assists - xA == 0 ~ "matching"
  ))

player_xG <- player_xG %>%
  mutate(Pos_Neg_A = case_when(
    assists - xA > 0 ~ "positive",
    assists - xA < 0 ~ "negative",
    assists - xA == 0 ~ "neutral"
  ))

performance_A <- reactive({
  player_xG$Performance_A[player_xG$player_name == input$player]
})

pos_neg_A <- reactive({
  player_xG$Pos_Neg_A[player_xG$player_name == input$player]
})
```

`r renderUI({HTML(paste("<b>", team(), "</b>"))})` have won `r renderText({wins()})` matches in the 2021/2022 Premier League season, drawing `r renderText({draws()})` and losing `r renderText({losses()})`.

Their home win percentage of `r renderText({label_percent()(home_win())})` ranks `r renderText({home_win_ranking()})` in the league, and their away win percentage of `r renderText({label_percent()(away_win())})` ranks at `r renderText({away_win_ranking()})`.

`r renderUI({HTML(paste("<b>", input$player, "</b>"))})` has scored `r renderText({goals()})` goals in the 2021/2022 season from an xG of `r renderText({round(xG(),2)})`, a `r renderText({pos_neg()})` delta of `r renderText({round(xG_delta(),2)})`. This means that he is `r renderText({performance()})` his xG this season.

The `r renderUI({HTML(paste("<b>", team(), "</b>"))})` player also has an xA of `r renderText({round(xA(),2)})`, from which he has recorded `r renderText({assists()})` assists, a `r renderText({pos_neg_A()})` delta of `r renderText({round(xA_delta(),2)})`. He is `r renderText({performance_A()})` his xA this season.

***

```{r, echo = FALSE}
# This code was sourced from https://github.com/FCrSTATS/StatsBomb_WomensData
# Create shot map
create_ShotMap <- function(grass_colour, line_colour, background_colour, goal_colour){

  theme_blankPitch = function(size=12) {
    theme(
      #axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      #axis.ticks.y=element_text(size=size),
      #   axis.ticks=element_blank(),
      axis.ticks.length=unit(0, "lines"),
      #axis.ticks.margin=unit(0, "lines"),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.background=element_rect(fill=background_colour, colour=NA),
      legend.key=element_rect(colour=background_colour,fill=background_colour),
      legend.key.size=unit(1.2, "lines"),
      legend.text=element_text(size=size),
      legend.title=element_text(size=size, face="bold",hjust=0),
      strip.background = element_rect(colour = background_colour, fill = background_colour, 
                                      size = .5),
      panel.background=element_rect(fill=background_colour,colour=background_colour),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      panel.spacing=element_blank(),
      plot.background=element_blank(),
      plot.margin=unit(c(0, 0, 0, 0), "lines"),
      plot.title=element_text(size=size*1.2),
      strip.text.y=element_text(colour=background_colour,size=size,angle=270),
      strip.text.x=element_text(size=size*1))}

  ymin <- 0 # minimum width
  ymax <- 100 # maximum width
  xmin <- 50 # minimum length
  xmax <- 100 # maximum length

  # Defining features along the length
  boxEdgeOff <- 83
  sixYardOff <- 94
  penSpotOff <- 88
  halfwayline <- 50

  # Defining features along the width
  boxEdgeLeft <- 18
  boxEdgeRight <- 82
  sixYardLeft <- 35
  sixYardRight <- 65
  goalPostLeft <- 46
  goalPostRight <- 54
  CentreSpot <- 50

  # other dimensions
  centreCirle_d <- 20

  # define the circle function
  circleFun <- function(center = c(0,0),diameter = 1, npoints = 100){
    r = diameter / 2
    tt <- seq(0,2*pi,length.out = npoints)
    xx <- center[1] + r * cos(tt)
    yy <- center[2] + r * sin(tt)
    return(data.frame(x = xx, y = yy))
  }
  
  #### create leftD arc ####
  dArc <- circleFun(c((50),(penSpotOff)),centreCirle_d,npoints = 1000)
  ## remove part that is in the box
  dArc <- dArc[which(dArc$y <= (boxEdgeOff)),]

  ## initiate the plot, set some boundaries to the plot
  p <- ggplot() + xlim(c(ymin,ymax)) + ylim(c(xmin,xmax)) +
    # add the theme
    theme_blankPitch() +
    # add the base rectangle of the pitch
    geom_rect(aes(xmin=ymin, xmax=ymax, ymin=xmin, ymax=xmax), fill = grass_colour, 
              colour = line_colour) +
    # add the 18 yard box offensive
    geom_rect(aes(xmin=boxEdgeLeft, xmax=boxEdgeRight, ymin=boxEdgeOff, ymax=xmax), 
              fill = grass_colour, colour = line_colour) +
    # add the six yard box offensive
    geom_rect(aes(xmin=sixYardLeft, xmax=sixYardRight, ymin=sixYardOff, ymax=xmax), 
              fill = grass_colour, colour = line_colour) +
    # add the arc circle
    geom_path(data=dArc, aes(x=x,y=y), colour = line_colour) +
    # add the goal offensive
    geom_segment(aes(x = goalPostLeft, y = xmax, xend = goalPostRight, yend = xmax),
                 colour = goal_colour, size = 1)

  return(p)

}
```

```{r, echo=FALSE}
# Spider/radar chart
spider_data = reactive({
  player_xG %>%
    filter(player_name == input$player) %>%
    select(player_name, xG90, xA90, npxG90, xgChain90, xgBuildup90) %>%
    column_to_rownames('player_name')
})

# Add the maximum, minimum row value
final_data <- reactive({
  rbind(rep(max(spider_data()), 5), rep(0, 5), spider_data())
})

source("shot_data.R")
shots_data <- reactive({(
  if (player_xG$goals[player_xG$player_name == input$player] == 0) {
    a = data.frame()
  } else {
    a = shot_data(input$player, player_xG)
  })
  return(a)
})

# Filter by 2021/2022 season and by goals
season_goals <- reactive({(
  if (nrow(shots_data()) == 0) {
    b = data.frame()
  } else {
    b = shots_data()[shots_data()$year == "2021" & shots_data()$result == "Goal",]
  })
  return(b)
})


# Separate out penalties
penalties <- reactive({(
  if (nrow(season_goals()[season_goals()$situation == "Penalty",]) == 0) {
    c = data.frame()
  } else {
    c = season_goals()[season_goals()$situation == "Penalty",]
  })
  return(c)
})

## Plot the base shot map
p <- create_ShotMap("#538032", "#ffffff", "#538032", "#000000")

gg_point <- reactive({(
  # For players who didn't score a single goal
  if (isempty(season_goals())) {
    d = p + 
  geom_text(aes(x = 2, y=59,label = input$player), hjust=0, vjust=0.5, size = 7, colour = "#ffffff")+
  geom_text(aes(x = 2, y=56,label = paste0("Expected Goals (xG): ",round(xG(),2))), hjust=0, vjust = 0.5, size = 5) +
    geom_text(aes(x = 2, y=54, label = paste0("Actual Goals: ", round(goals(), 2))), hjust=0, vjust=0.5, size=5) +
    geom_text(aes(x = 2, y=52, label = paste0("Delta: ", round(xG_delta(),2))), hjust=0, vjust=0.5, size=5) +
      scale_size(name = "xG")
  # For players who scored at least one goal but no penalties  
  } else if (!isempty(season_goals()) && isempty(penalties())) {
    d = p + geom_point_interactive(data = season_goals(), aes(x=(1-Y)*100, y=X*100, size=round(xG,1), tooltip=paste0("vs. ", Opponent, "\n xG: ", round(xG,2), "\n Last action: ", lastAction)), colour = "#ffffff") +
  geom_text(aes(x = 2, y=59,label = input$player), hjust=0, vjust=0.5, size = 7, colour = "#ffffff")+
  geom_text(aes(x = 2, y=56,label = paste0("Expected Goals (xG): ",round(xG(),2))), hjust=0, vjust = 0.5, size = 5) +
    geom_text(aes(x = 2, y=54, label = paste0("Actual Goals: ", round(goals(), 2))), hjust=0, vjust=0.5, size=5) +
    geom_text(aes(x = 2, y=52, label = paste0("Delta: ", round(xG_delta(),2))), hjust=0, vjust=0.5, size=5)  +
      scale_size(name = "xG")
  # For players who scored at least one goal and at least one penalty  
  } else {
    d = p + geom_point_interactive(data = season_goals(), aes(x=(1-Y)*100, y=X*100, size=round(xG,1), tooltip=paste0("vs. ", Opponent, "\n xG: ", round(xG,2), "\n Last action: ", lastAction)), colour = "#ffffff") +
    
    geom_point_interactive(data = penalties(), aes(x=(1-Y)*100,y=X*100,size=xG,
                                                   tooltip=paste0("Goals scored from penalties: ", nrow(penalties()))),colour = "#000000") + 
    
  geom_text(aes(x = 2, y=59,label = input$player), hjust=0, vjust=0.5, size = 7, colour = "#ffffff")+
  geom_text(aes(x = 2, y=56,label = paste0("Expected Goals (xG): ",round(xG(),2))), hjust=0, vjust = 0.5, size = 5) +
    geom_text(aes(x = 2, y=54, label = paste0("Actual Goals: ", round(goals(), 2))), hjust=0, vjust=0.5, size=5) +
    geom_text(aes(x = 2, y=52, label = paste0("Delta: ", round(xG_delta(),2))), hjust=0, vjust=0.5, size=5)  +
      scale_size(name = "xG")
  }
  
  
)
  return(d)
})

fluidRow(
  
  column(h4("Shot position of goals scored"),
    p("When hovering over a point, the following is displayed:", br(), br(
                      "1. Opponent against which goal was scored", br(), br(
                        "2. xG associated with the shot", br(), br(
                          "3. The action which led to the shot being taken", br(), br(
                            "Goals scored from penalties are coloured in black"
                          )))),
                    style="color:black;text-align:left"),
                  width=6,style="background-color:lavender;border-radius: 10px;height: 300px"),
  
  column(h4("Visualising player performance"),
    p("Explanation of terms:", br(), br(
                    "1. xG90/xA90 = Expected goals/assists per 90 minutes", br(), br(
                      "2. npxG90 = Non-penalty expected goals per 90 minutes", br(), br(
                        "3. xgChain90 = Total xG of every possession the player is involved in
                        per 90 minutes", br(), br(
                          "4. xgBuildup90 = Total xG of every possession the player is involved
                          in without key passes and shots per 90 minutes"
                        )))),
                    style="color:black;text-align:left"),
                  width=6,style="background-color:lavender;border-radius: 10px;height: 300px")
)

fluidRow()

fluidRow(
  
  column(6,
         renderggiraph(
           girafe(ggobj = gg_point())
 )),
  
  column(6,
         renderPlot({
           radarchart(final_data(),
                      seg = 5)}))
  
)
```